### 2017-07-06 lafaySimulatedScenePerception.tex
ml proof read 1

\Englishabstract{This paper introduces a new experimental protocol to study mental representations of urban soundscapes through a simulation process. Subjects are asked to create a full soundscape by means of a software tool dedicated to sound edition, coupled with a structured sound data set. This paradigm is used to characterize urban sound environment representations by analyzing the sound classes that were used to simulate the auditory scenes. A rating experiment of the soundscape pleasantness using a 7-points Likert scale is conducted to further refine the analysis of the simulated urban acoustic scenes. Results show that 1) a semantic characterization in terms of presence~/ absence of sound sources is an effective way to characterize urban soundscapes pleasantness, and 2) physical descriptors computed for specific sound sources better characterize the appraisal than global descriptors.}
\Germanorfrenchabstract{}

@@

First, a subject can be asked to describe a given sound  following a given perceptive scale (\eg~calmness, pleasantness) \cite{axelsson2005soundscape,davies2013perception,cain2013development}. The amount of information that can be gathered strongly depends on the very nature of the available stimuli, be they recorded sound scenes as part of a within laboratory experiment, or \emph{in situ}. That being said, as those stimuli can be recorded and analysed, it allow the experimenter to gather a coarse grain physical description of the stimuli.

@@

Second, a subject can be asked to describe a given sound environment \cite{guastavino2006ideal, dubois2006cognitive}. A large amount of quantitative and semantic information is then collected about the representation of this type of sound environment that the subject has. Unfortunately, without any reference to sound data, this representation cannot be characterized physically.

@@

For several reasons that are detailed in this paper, the use of a soundscape simulator such as the one proposed in this article can lead to very interesting outcomes as it allows the experimenter to separately study  the influence of the sounds sources that compose a sound environment of interest. Indeed, with such a material, not only the type of sound sources is available for study, but also the exact level and audio waveform for each source and also the structural characteristics of the scene, that is the temporal distribution of the events.

@@

To demonstrate the potential of the proposed approach in its ability to question the human perception of the sound environment, we study in this paper the notion of perceptual pleasantness of urban environmental scenes. Results and outcomes of a series of three experiments that build on the use of the simulator are studied in order to better comprehend how different sound sources typically present in a urban scene impact pleasantness:

@@

To the best of our knowledge, only Bruce~\al \cite{bruce2009development,bruce2014effects} considered the use of a simulator to question soundscape perception. They propose a tool that allows the user to modify a given soundscape by adding or removing specific sound sources, by changing the acoustic level of the sources as well as their spatial location. The authors shows that the addition or removal of the sources globally follows more social or semantic matters than their acoustical characteristics. A lack of diversity is nonetheless mentioned by the authors that limit the outcomes of the study.

@@

In our approach, the simulator developed for this study\footnote{simScene, available online \url{http://soundthings.org/research/}} gives a monophonic representation of the scene, with the added benefit of a wider range of available sound sources and scheduling parameters in order to provide outputs which are, as much as possible, expressive and useful for analysis.

@@

\emph{Simscene} is designed as a simplified audio sequencer, with sequencing parameters specifically chosen for the generation of realistic soundscapes. To do so, the user first selects a sound source using a non verbal selection interface presented in Section~\ref{sec:simscene_ssf}. A track is then created for this sounds source within the simulator interface. The user can then manipulate some parameters detailed in Section~\ref{sec:simscene_parametre} to control the time and magnitude distribution of the occurrences of the sounds source. Text fields are also available for the user to 1) name each track, 2) name the entire scene, 3) to provide some comments about the simulated scene.

@@

The chosen typology is established based on the category/classes of sounds found while reviewing 16 articles or thesis manuscripts \cite{maffiolo_caracterisation_1999,raimbault2002simulation,guastavino_etude_2003,defreville2004aactivity,raimbault2005urban,dubois2006cognitive,devergie_relations_2006,guastavino2006ideal,niessen2010categories,maffiolo_caracterisation_1999,beaumont2004pertinence,polack2008perceptive,leobon_analyse_1986,brown2011towards} that study how humans discriminate different kind of urban soundscapes.

@@

Literature usually distinguish between:
\begin{itemize}
\item {a sound event}: an isolated sound, sparsely distributed whose acoustical characteristics may change with respect to time
\item {a sound texture}: an isolated sound of long duration whose acoustical characteristics are stable with respect to time \cite{saint1995classification}.
\end{itemize}

The simulation tool \emph{Simscene} follows this distinction and consider two distinct sound databases: one with the classes of events only and the other with classes of textures. These two types have specific simulation procedures.

@@

The event sequences are composed of salient and identifiable events such as \emph{car start}, \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On contrary, the amorphous sequences are environments where distinct events can hardly be found, such as \emph{traffic hubhub}. They build upon an holistic analysis based on global acoustic features.

@@

The sound classes are organized hierarchically (\cf~Figure~\ref{fig:orgDb}) according to a structure that is close to the one of the vertical axis of the categorical organization proposed by E. Rosch \cite{rosch1978cognition}. The lower the level of abstraction, the more precise the description of the class and the more perceptually similar the sound sources. For classes with high level of abstraction with sub classes, its collection of samples is the union of the collections of the sub classes.

@@

Accounting for the perceptual matters detailed previously, two taxonomies are built, one for event sounds and one for texture sounds. Four levels of abstraction are considered from the more generic one (level 0), to the more specific classe (level 3), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

Original sounds have been recorded using a shotgun microphone \emph{AT8035}\footnote{\emph{Micro canon AT8035}: \url{eu.audio-technica.com}} plugged to a \emph{ZOOM H4n}\footnote{\emph{ZOOM H4n} : \url{www.zoom.co.jp/english/products/h4n}} recorder. The use of such a microphone allowed us to isolate as much as possible sound events of interest from the urban background. It also allowed us to avoid dominant sound sources while recording texture sounds by aiming the microphone to distant areas with no dominant sound sources.

>>>>

\Englishabstract{This paper introduces a new experimental protocol to study mental representations of urban soundscapes through a simulation process. Subjects are asked to create a full soundscape by means of a dedicated software tool, coupled with a structured sound data set. This paradigm is used to characterize urban sound environment representations by analyzing the sound classes that were used to simulate the auditory scenes. A rating experiment of the soundscape pleasantness using a 7-points Likert scale is conducted to further refine the analysis of the simulated urban acoustic scenes. Results show that 1) a semantic characterization in terms of presence~/ absence of sound sources is an effective way to characterize urban soundscapes pleasantness, and 2) physical descriptors computed for specific sound sources better characterize the appraisal than global descriptors.}
\Germanorfrenchabstract{}

@@

First, a subject can be asked to describe a given sound following a given perceptive scale (\eg~calmness, pleasantness) \cite{axelsson2005soundscape,davies2013perception,cain2013development}. The amount of information that can be gathered strongly depends on the very nature of the available stimuli, be they recorded sound scenes as part of a within laboratory experiment, or \emph{in situ}. That being said, those stimuli respectively are or can be recorded and analysed. It thus allows the experimenter to gather a coarse grain physical description of the stimuli.

@@

Second, a subject can be asked to describe a given sound environment \cite{guastavino2006ideal, dubois2006cognitive}. A large amount of quantitative and semantic information is then collected about the representation of this type of sound environment that the subject has. Unfortunately, without any reference to sound data, this representation can hardly be characterized physically.

@@

For several reasons that are detailed in this paper, the use of a soundscape simulator such as the one proposed in this article can lead to interesting outcomes as it allows the experimenter to separately study  the influence of the sounds sources that compose a sound environment of interest. With such a material, not only the type of sound sources is indeed available for study, but also the exact level and audio waveform for each source together with the structural characteristics of the scene, that is the temporal distribution of the events.

@@

To demonstrate the potential of the proposed approach in its ability to question the human perception of the sound environment, we study in this paper the notion of perceptual pleasantness of urban environmental scenes. Results and outcomes of a series of three experiments that foster on the use of the simulator are studied in order to better comprehend how different sound sources typically present in a urban scene impact pleasantness:

@@

To the best of our knowledge, only Bruce~\al \cite{bruce2009development,bruce2014effects} considered the use of a simulator to question soundscape perception. They propose a tool that allows the user to modify a given soundscape by adding or removing specific sound sources, by changing the acoustic level of the sources as well as their spatial location. The authors shows that the addition or removal of the sources globally follows more social or semantic matters than their acoustical characteristics. A lack of diversity is nonetheless mentioned by the authors that limit the strenght of the outcomes given in this study.

@@

In our approach, the simulator developed for this study\footnote{simScene, available online \url{https://bitbucket.org/mlagrange/simscene}} gives a monophonic representation of the scene, with the added benefit of a wider range of available sound sources and scheduling parameters in order to provide outputs which are both expressive and useful for analysis.

@@

\emph{Simscene} is designed as a simplified audio sequencer, with sequencing parameters specifically chosen for the generation of realistic soundscapes. To do so, the user first selects a sound source using a non verbal selection interface presented in Section~\ref{sec:simscene_ssf}. A track is then created for this sound source within the simulator interface. The user can then manipulate some parameters detailed in Section~\ref{sec:simscene_parametre} to control the time and magnitude distribution of the occurrences of the sounds source. Text fields are also available for the user to 1) name each track, 2) name the entire scene, 3) to provide some comments about the simulated scene.

@@

The chosen typology is established based on the category/classes of sounds found while reviewing several articles or thesis manuscripts \cite{maffiolo_caracterisation_1999,raimbault2002simulation,guastavino_etude_2003,defreville2004aactivity,raimbault2005urban,dubois2006cognitive,devergie_relations_2006,guastavino2006ideal,niessen2010categories,maffiolo_caracterisation_1999,beaumont2004pertinence,polack2008perceptive,leobon_analyse_1986,brown2011towards} that study how humans discriminate different kind of urban soundscapes.

@@

The literature usually distinguish between:
\begin{itemize}
\item {a sound event}: an isolated sound, sparsely distributed whose acoustical characteristics may change with respect to time
\item {a sound texture}: an isolated sound of long duration whose acoustical characteristics are stable with respect to time \cite{saint1995classification}.
\end{itemize}

The simulation tool \emph{Simscene} follows this distinction and consider two distinct sound databases: one with the classes of events only and the other with classes of textures only. These two types also have specific scheduling procedures during the sumulation process.

@@

The event sequences are composed of salient events that can easily be recognized such as \emph{car start} or \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On contrary, the amorphous sequences are sound environments where distinct events can hardly be found, such as \emph{traffic hubhub}. They build upon a holistic analysis based on global acoustic features.

@@

The sound classes are organized hierarchically (\cf~Figure~\ref{fig:orgDb}) according to a structure that is close to the one of the vertical axis of the categorical organization proposed by E. Rosch \cite{rosch1978cognition}. The lower the level of abstraction, the more precise the description of the class and the more perceptually similar the sound sources is. For classes with high level of abstraction that have sub classes, its collection of samples is the union of the collections of the sub classes.

@@

Accounting for the perceptual matters detailed previously, two taxonomies are built, one for event sounds and one for texture sounds. Four levels of abstraction are considered from the more generic one (level 0), to the more specific classes (level 3), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

Original sounds have been recorded using a shotgun microphone \emph{AT8035}\footnote{\emph{Micro canon AT8035}: \url{eu.audio-technica.com}} plugged to a \emph{ZOOM H4n}\footnote{\emph{ZOOM H4n} : \url{www.zoom.co.jp/english/products/h4n}} recorder. The use of such a microphone allowed us to isolate as much as possible sound events of interest from the urban background. It also allowed us to avoid dominant sound sources while recording texture sounds by targeting distant areas with no dominant sound sources.
